---
title: "STAT 345 Midterm"
author: "Brandon Miller"
date: '2022-03-25'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
#loading in packages and installing necessary data
library(tidyverse)
library(devtools)
install_github("abresler/nbastatR")
library(nbastatR)
library(ggplot2)
#changing the size of connection (this came up every time I installed the data in so I had to do this in order to fix it)
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)
#read in 20 years of data
shots_2002 <- teams_shots(teams="Milwaukee Bucks", seasons = 2002)
shots_2003 <- teams_shots(teams="Milwaukee Bucks", seasons = 2003)
shots_2004 <- teams_shots(teams="Milwaukee Bucks", seasons = 2004)
shots_2005 <- teams_shots(teams="Milwaukee Bucks", seasons = 2005)
shots_2006 <- teams_shots(teams="Milwaukee Bucks", seasons = 2006)
shots_2007 <- teams_shots(teams="Milwaukee Bucks", seasons = 2007)
shots_2008 <- teams_shots(teams="Milwaukee Bucks", seasons = 2008)
shots_2009 <- teams_shots(teams="Milwaukee Bucks", seasons = 2009)
shots_2010 <- teams_shots(teams="Milwaukee Bucks", seasons = 2010)
shots_2011 <- teams_shots(teams="Milwaukee Bucks", seasons = 2011)
shots_2012 <- teams_shots(teams="Milwaukee Bucks", seasons = 2012)
shots_2013 <- teams_shots(teams="Milwaukee Bucks", seasons = 2013)
shots_2014 <- teams_shots(teams="Milwaukee Bucks", seasons = 2014)
shots_2015 <- teams_shots(teams="Milwaukee Bucks", seasons = 2015)
shots_2016 <- teams_shots(teams="Milwaukee Bucks", seasons = 2016)
shots_2017 <- teams_shots(teams="Milwaukee Bucks", seasons = 2017)
shots_2018 <- teams_shots(teams="Milwaukee Bucks", seasons = 2018)
shots_2019 <- teams_shots(teams="Milwaukee Bucks", seasons = 2019)
shots_2020 <- teams_shots(teams="Milwaukee Bucks", seasons = 2020)
shots_2021 <- teams_shots(teams="Milwaukee Bucks", seasons = 2021)
#combine years
shots <- rbind(shots_2002, shots_2003, shots_2004, shots_2005, shots_2006, shots_2007, shots_2008, shots_2009, shots_2010, shots_2011, shots_2012, shots_2013, shots_2014, shots_2015, shots_2016, shots_2017, shots_2018, shots_2019, shots_2020, shots_2021)
#graphed new years of shots
shots_chart <- ggplot(data=shots, aes(x=locationX, y=locationY)) + geom_point() + xlim(-300, 300) + ylim(-100, 500)
```

```{r}
#creating the basketball court
library(ggplot2)
 
court <- ggplot(data=data.frame(x=1,y=1),aes(x*10,y*10))+
   ###outside box:
geom_path(data=data.frame(x=c(-27,-27,27,27,-27),y=c(-4,50,50,-4,-4)))+
   ###key:
geom_path(data=data.frame(x=c(-8,-8,8,8,-8),y=c(-4,14,14,-4,-4)))+
   ###box inside the key:
geom_path(data=data.frame(x=c(-6,-6,6,6,-6),y=c(-4,14,14,-4,-4)))+
   ###restricted area semicircle:
geom_path(data=data.frame(x=c(-4000:(-1)/1000,1:4000/1000),y=c(0+sqrt(4^2-c(-4000:(-1)/1000,1:4000/1000)^2))),aes(x=10*x,y=11*y))+
   ###halfcourt semicircle:
geom_path(data=data.frame(x=c(-6000:(-1)/1000,1:6000/1000),y=c(50-sqrt(6^2-c(-6000:(-1)/1000,1:6000/1000)^2))),aes(x=10*x,y=10*y))+
   ###rim:
geom_path(data=data.frame(x=c(-750:(-1)/1000,1:750/1000,750:1/1000,-1:-750/1000),y=c(c(-3+sqrt(0.75^2-c(-750:(-1)/1000,1:750/1000)^2)),c(-3-sqrt(0.75^2-c(750:1/1000,-1:-750/1000)^2)))),aes(x=10*x,y=10*y))+
   ###backboard:
geom_path(data=data.frame(x=c(-3,3),y=c(-4,-4)),lineend='butt')+
   ###three-point line:
geom_path(data=data.frame(x=c(-22,-22,-22000:(-1)/1000,1:22000/1000,22,22),y=c(-5,169/12,5.25+sqrt(23.75^2-c(-22000:(-1)/1000,1:22000/1000)^2),169/12,-5)),aes(x=10*x,y=8*y))+
   ###fix aspect ratio to 1:1
coord_fixed()

```
```{r message=FALSE, warning=FALSE}
#load packages for gganimate
library(gganimate)
library(tidyverse)
library(gapminder) #for some data
library(gifski) #creation of gifs, could also use library(png)
library(lubridate) #working with time variables
library(maps) #for mapping with ggplot
library(mapdata) #for mapping with ggplot
library(transformr) #for changing chloropleths across time
library(ggthemes) #extra color options
library(viridis)
#creating shot chart of data
shotchart <- court + transition_time(yearSeason)+ labs(title="Year: {frame_time}", x="", y="")+theme_bw()+geom_hex(data=shots, aes(x=locationX, y=locationY), bins=25, alpha=0.8)+scale_fill_viridis(option = "C", trans="log")
#animate data
animate(shotchart, fps=1, nframes=20)
```

- The graphic shown is the progression of NBA shots for the Milwaukee Bucks from 2002 to 2021.
- We can see that as the years progression shots from behind the 3 point line increase as the shot density increases there. 
- We can also see the extreme that even more shots are being taken from beyond half court as years progress. 

Report:
The above shot chart is given from the data given by the nbastatR package and uses the shots of the Milwaukee Bucks from 2002 to 2021. My goal was to see the progression of shots over the past couple of decades since people have been saying that players have started to shoot from farther away than 20 years ago. I needed to first create the court as well as plotting the shots of each year on the court. I proceeded to create a graphic where I animated in which at each frame, it shows the shots from year to year. Lastly I wanted to show the density of shots is certain areas so I used hexagon bins and used a log scale to distinguish the difference in shot density from each part of the court. Based on the graphic, we can see that as the years progress, shots start to expanded to beyond the 3 point line as well as more shots are taken beyond half-court. This shows that the hypothesis that players have started to shot from further away than 20 years is correct and should be explored more into the reason why this is the case. 
   
```{r}
#grab draft data
drafts <- drafts(draft_years = 1948:2021)
#grab mvp award data
mvp <- bref_awards(awards="Most Valuable Player")
#join data sets by name of player
drafts_mvp <- left_join(drafts, mvp, by="namePlayer") %>% filter(slugAward=="mvp") %>% select("namePlayer", "numberPickOverall")
#fixed typo in data frame
drafts_mvp[drafts_mvp==0] <- 3
#finding the number of mvps at each draft pick
drafts_mvp_count <- drafts_mvp %>% group_by(numberPickOverall) %>% summarise(count=n())
#join more data
drafts_mvp_n <- left_join(drafts_mvp, drafts_mvp_count, by="numberPickOverall")
#plotting previous information in bar plot
ggplot(data=drafts_mvp_n, aes(x=numberPickOverall, fill=namePlayer)) + geom_bar()+labs(title="Total Number of MVPs for Each Overall Draft Selection", x="Overall Pick", y="Number of MVPs")
```

